---
title: "The Hobbit Text Mining"
author: "Fisher Ankney"
date: "1/25/2018"
output: html_document
---
After recently finishing J.R.R. Tolkein's classic, *The Hobbit*, I decided to try my hand at text mining in R. Now, I've never worked worked with large strings of text in R before, so I decided to look around the internet for a little bit of help and inspiration. After reading about [Text Mining with R](https://www.tidytextmining.com/),
R-Bloggers Metallica lyric analysis [post](https://www.r-bloggers.com/hardwired-for-tidy-text/), and 
Data Camp's text mining blog [post](https://www.datacamp.com/community/blog/text-mining-in-r-and-python-tips), I began to understand the challenge that awaited, and what I wanted to explore in Tolkein's work. 

I decided to break up my Hobbit text project into three parts - A traditional word frequency analysis, An individual character assessment, and finally various measurements sentiment throughout the novel. If you want to learn more about text mining in R, or if you're just interested in hobbits I hope you find this investigation helpful. [Here](https://archive.org/stream/TheHobbitByJ.R.RTolkien/The%20Hobbit%20by%20J.R.R%20Tolkien_djvu.txt) is the copy of the hobbit that I'm using if you want to follow along. 

First things first, let's load up the libraries we'll be using and tidy up the dataset. 

<br  />

Load Libraries
```{r, message=FALSE}
library('tidyverse')    # used for EVERYTHING
library('stringr')      # used for string (text) manipulation
library('tidytext')     # used for text mining
library('wordcloud')    # used for creating cool text visualizations
```


<br  />

Load Data
```{r}
the_hobbit <- as.tibble(read_file("input/hobbit_body.txt")) 
```

<br  />

Tidy Data 
```{r, message=FALSE}
tidy_hobbit <- the_hobbit %>%      # move the_hobbit text into tidy_hobbit
  unnest_tokens(word, value) %>%   # each word is given a cell in tibble
  anti_join(stop_words)            # take out meaningless words lke a, to, and it
```

<br  />

So here's the first look at our Hobbit text. A quick note: I cleaned up the hobbit text to start at Chapter 1 before I saved it as an input file. *head(the_hobbit)* reveals a 1x1 table with the entire novel stored as a single string in the tibble. This formatting has some uses so I'll keep a copy of it, but the tidy text formatted *tidy_hobbit* has more uses.

*tidy_hobbit* has cut out some words - in, a, in, the, there, and so on. That's the *anti_join(stop_words)* at work. *anti_join()* is a dplyr function we loaded from the tidyverse, it's job is to return all rows from x where there are not matching values in y. In this case x is *the_hobbit*, y is stop_words, a tibble from the tidytext library. Also notice that now each word has it's own row in the tibble, which is now length over 30,000 elements long.

<br  />

View The Data
```{r}
head(the_hobbit)
```

```{r}
head(tidy_hobbit)
```

<br  />

### Word Frequency

<br  />

How many words are in *The Hobbit*?
```{r}
str_count(the_hobbit, " ") + 1
```

<br  />

How many sentences?
```{r}
str_count(the_hobbit, "[.!?]")
```

<br  />

How long should the book take to read?
```{r}
((str_count(the_hobbit, " ") + 1) / 200) / 60 # number of words / 200 words-per-minute / 60 minutes
```

<br  />

#### Frequency Plot

What are the most common meaningful words in *The Hobbit*? To find out, let's create a frequency plot. These overturned bar graphs are some of my favorite to create.

```{r}
tidy_hobbit %>%
  count(word, sort = TRUE) %>%          # count words
  filter(n > 80) %>%                    # only show words that have been used 80+ times
  mutate(word = reorder(word, n)) %>%   # sorts words by decending occurance
  ggplot(aes(word, n)) +                # plot the words vs number of occurances
    geom_col(fill = "darkslateblue") +  # blue bars
  xlab("Word") +                        # labels
  ylab("Number of Occurances") + 
  labs(title = "Most Common Meaningful Words in 'The Hobbit'") +
  coord_flip() +                        # sideways box-plot for aestetic purposes
  theme_minimal()                       # I don't like the gray default theme
```

The most common words in *The Hobbit* are Bilbo, dwarves, Thorin, time, and Gandalf. The only non-character word is time, but I bet the word time is pretty common in most novels. I'm suprised to see the word "don't" on the list, I would have thought it was classified as a stop word. There isn't much to be suprised about on this list if you've read the book. If you haven't read the book, the list gives you a pretty good idea of what to expect!

<br  />

#### Word Cloud
```{r}
hobbit_cloud <- tidy_hobbit %>%
  count(word) %>%                                      # provide a count of each word
  with(wordcloud(word, n, max.words = 100,             # using the wordcloud library, plot word, according to n (number of count), with 100 maximum words)
                 color = brewer.pal(n = 8,"Set1")))    # use RColorBrwer pallete 'Set1'
```


The top 100 words used in The Hobbit are all jumbled together in this word cloud. It's basically the same idea as the frequency plot, except it gives less information in a more artistic way. I'm not sure what's with the color coding, I would have thought all 8 pallete colors would be used, not just five. The size of the words reflect their frequency though. 

<br  />

## Character Analysis

Now let's dig into the main characters of the novel. Who are the most important characters and where do they appear most in the book?

<br  />

#### Main Character Count
```{r}
main_chr <- as.tibble(c("bilbo", "thorin", "gandalf", "smaug", "gollum", "bard")) %>%
  unnest_tokens(word, value)

colnames(main_chr) <- "word"

main_chr_num <- tidy_hobbit %>% 
  semi_join(main_chr, "word") %>%
  count(word, sort = TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) + 
    geom_col(fill = "darkslateblue")
main_chr_num
```

As we saw before, Bilbo is not only the most important character out of the six I've ranked, bilbo is the most common word of the entire book. Life is good being the protagonist! Thorin the dwarf king is mentioned some 250 times, and Gandalf is mentioned under 200 times. I would have thought Gandalf played a bigger role than Thorin, but I guess the stats don't lie. Ol' Gollum, Smaug, and Bard are pretty minor characters.

<br  />

```{r}

tidy_hobbit$linenumber <- 1:nrow(tidy_hobbit)

main_chr_pres <- tidy_hobbit %>% 
  semi_join(main_chr, "word") %>%
  ggplot() + 
  geom_point(
    aes(word, linenumber)
  ) + 
  labs(title = "Main Character Apperances") +
  xlab("Character") + 
  ylab("")+
  coord_flip() + 
  theme_minimal() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
main_chr_pres
  
```

By the looks of it, Bilbo is mentioned every other sentence! There's two small gaps in Bilbos presence near the end of the book where it seems Bard takes the center stage. Devilish old Smaug is breifly mentioned in the first half of the book, something along the lines of "oh there's also a dragon guarding the gold but don't worry about it". Gandalf has an interesting timeline, he has three solid apperances in the tale, sitting out for the majority of the story-climatic burglary. 

<br  />

#### Dwarf Character Count

```{r}
dwarf_chr <- as.tibble(c("dwalin", "balin", "kili", "fili", "dori", "nori", "ori", "oin", "gloin", "bifur", "bofur", "bombur", "thorin")) %>%
  unnest_tokens(word, value)

colnames(main_chr) <- "word"

dwarf_chr_num <- tidy_hobbit %>% 
  semi_join(dwarf_chr, "word") %>%
  count(word, sort = TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) + 
    geom_col(fill = "darkslateblue")
dwarf_chr_num
```

Poor Ori. It seems like Oir, Nori, Bifur, Oin, Bofur, Dwalin and Gloin are nothing but background characters. Fili is mentioned noticably more than his twin Kili, and the roundest of the dwarves, Bombur is actually pretty popular in his comic relief. Balin is much more important than his brother Dwalin, which makes sense as he's the second eldest dwarf and often the leader of the bunch in Thorin's absence.

<br  />


```{r}
dwarf_chr_pres <- tidy_hobbit %>% 
  semi_join(dwarf_chr, "word") %>%
  ggplot() + 
  geom_point(
    aes(word, linenumber)
  ) + 
  labs(title = "Dwarve Apperances") +
  xlab("Dwarf") + 
  ylab("")+
  coord_flip() + 
  theme_minimal() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
dwarf_chr_pres
```


Ignoring Thorin, Balin, and Bombur, each of the other ten dwarves are almost always mentioned all together. Fili and Kili's presence is almost identical, it's hard to spot why Fili is apparently more prominent. 

<br  />

## Word Sentiment

Using the tidy text library we can analyze sentiment throughout the tale. Let's just jump in and start exploring! 

```{r, message=FALSE, warning=FALSE}
nrcjoy <- get_sentiments("nrc") %>%    # using the nrc sentiment package in the tidy text library
  filter(sentiment == "joy")           # using only words that have a sentiment of joy

happy_cloud <- tidy_hobbit %>%         # put the tidy_hobbit into a new variable
  inner_join(nrcjoy) %>%               # only keep words that match the nrcjoy variable 
  count(word) %>%                      # now count the number of times each word is used
  with(wordcloud(word, n, max.words = 100,  # create a word cloud based on this
                 color = brewer.pal(n = 8,"Dark2")))
happy_cloud
```

The top 100 words to describe joy in *The Hobbit*! I think this is the coolest visualization yet. So many of these adjectives just nail the adventurous spirit and comradery present in the book.


```{r}
joy_meter <- tidy_hobbit %>% 
  inner_join(get_sentiments("nrc")) %>% 
  filter(sentiment == "joy") %>%
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = n()) %>% 
  mutate(method = "NRC")

joy_meter %>%
  ggplot(aes(index, sentiment)) +
  geom_col(show.legend = FALSE,
           fill = "orange") + 
  xlab("Book Progression") + 
  ylab("Joy Meter") + 
  labs(title = "Words with Joyous Sentiment in 'The Hobbit'") +
  theme_minimal()
```



joy joy joy


```{r}
# Sad Words 
nrcsad <- get_sentiments("nrc") %>%    # from the tidytext
  filter(sentiment == "sadness")

sad_cloud <- tidy_hobbit %>%
  inner_join(nrcsad) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100,
                 color = brewer.pal(n = 8,"Dark2")))
sad_cloud

# Frequency of sadness throught the book
sad_meter <- tidy_hobbit %>% 
  inner_join(get_sentiments("nrc")) %>% 
  filter(sentiment == "sadness") %>%
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = n()) %>% 
  mutate(method = "NRC")

sad_meter %>%
  ggplot(aes(index, sentiment)) +
  geom_col(show.legend = FALSE, 
           fill = "lightsteelblue") + 
  xlab("Book Progression") + 
  ylab("Sad Meter") + 
  labs(title = "Words with Sad Sentiment in 'The Hobbit'") +
  theme_minimal() 
```
  
  
  
```{r}  
## Fearful Words
nrcfear <- get_sentiments("nrc") %>%   
  filter(sentiment == "fear")

fear_cloud <- tidy_hobbit %>%
  inner_join(nrcfear) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100,
                 color = brewer.pal(n = 8,"Dark2")))


# Frequency of sadness throught the book
fear_meter <- tidy_hobbit %>% 
  inner_join(get_sentiments("nrc")) %>% 
  filter(sentiment == "fear") %>%
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = n()) %>% 
  mutate(method = "NRC")

fear_meter %>%
  ggplot(aes(index, sentiment)) +
  geom_col(show.legend = FALSE, 
           fill = "Red4") + 
  xlab("Book Progression") + 
  ylab("Fear Meter") + 
  labs(title = "Words with Fear Sentiment in 'The Hobbit'") +
  theme_minimal()
```


```{r}
### Happy, Sad, and Fear all together now
ggplot() + 
  geom_smooth(aes(fear_meter$index,
                  fear_meter$sentiment, 
                  color = "Fear"),
              se = FALSE) + 
  geom_smooth(aes(joy_meter$index, 
                  joy_meter$sentiment, 
                  color = "Joy"),
              se = FALSE) + 
  geom_smooth(aes(sad_meter$index, 
                  sad_meter$sentiment, 
                  color = "Sadness"),
              se = FALSE) + 
  xlab("Book Progression Index") + 
  ylab("Sentiment Index") +
  labs(
    title = "Specific Sentiment in 'The Hobbit'"
  ) + 
  theme_minimal() + 
  scale_color_manual(
    name = "Sentiment",
    values = c("red4", "orange", "lightsteelblue")
  )
```

```{r}
# Overall Sentiment
hobbit_sent <- tidy_hobbit %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = sum(score)) %>% 
  mutate(method = "AFINN")

hobbit_sent[["sign"]] = ifelse(hobbit_sent[["sentiment"]] >=0, "positive", "negative")

hobbit_sent %>%
  ggplot() + 
  geom_col(aes(index, 
               sentiment,
               fill = sign)
           ) + 
  scale_fill_manual(name = "Sentiment",
                    values = c("positive" = "royalblue4", "negative" = "red4"),
                    guide = guide_legend(reverse=TRUE)) + 
  xlab("Book Progression Index") + 
  ylab("Sentiment Index") + 
  ggtitle("General AFINN Method Sentiment in 'The Hobbit'") + 
  theme_minimal()
```


the end
















