---
title: "The Hobbit Text Analysis"
author: "Fisher Ankney"
date: "1/25/2018"
output: html_document
---
After recently finishing J.R.R. Tolkein's classic, *The Hobbit*, I decided to try my hand at text mining in R. Now, I've never worked worked with large strings of text in R before, so I decided to look around the internet for a little bit of help and inspiration. After reading about [Text Mining with R](https://www.tidytextmining.com/),
R-Bloggers Metallica lyric analysis [post](https://www.r-bloggers.com/hardwired-for-tidy-text/), and 
Data Camp's text mining blog [post](https://www.datacamp.com/community/blog/text-mining-in-r-and-python-tips), I began to understand the challenge that awaited, and what I wanted to explore in Tolkein's work. 

I decided to break up my Hobbit text project into three parts - A traditional word frequency analysis, An individual character assessment, and finally various measurements sentiment throughout the novel. If you want to learn more about text mining in R, or if you're just interested in hobbits I hope you find this investigation helpful. [Here](https://archive.org/stream/TheHobbitByJ.R.RTolkien/The%20Hobbit%20by%20J.R.R%20Tolkien_djvu.txt) is the copy of the hobbit that I'm using if you want to follow along. 

First things first, let's load up the libraries we'll be using and tidy up the dataset. 

<br  />

Load Libraries
```{r, message=FALSE}
library('tidyverse')    # used for EVERYTHING
library('stringr')      # used for string (text) manipulation
library('tidytext')     # used for text mining
library('wordcloud')    # used for creating cool text visualizations
```


<br  />

Load Data
```{r}
the_hobbit <- as.tibble(read_file("input/hobbit_body.txt")) 
```

<br  />

Tidy Data 
```{r, message=FALSE}
tidy_hobbit <- the_hobbit %>%      # move the_hobbit text into tidy_hobbit
  unnest_tokens(word, value) %>%   # each word is given a cell in tibble
  anti_join(stop_words)            # take out meaningless words lke a, to, and it
```

<br  />

So here's the first look at our Hobbit text. A quick note: I cleaned up the hobbit text to start at Chapter 1 before I saved it as an input file. *head(the_hobbit)* reveals a 1x1 table with the entire novel stored as a single string in the tibble. This formatting has some uses so I'll keep a copy of it, but the tidy text formatted *tidy_hobbit* has more uses.

*tidy_hobbit* has cut out some words - in, a, in, the, there, and so on. That's the *anti_join(stop_words)* at work. *anti_join()* is a dplyr function we loaded from the tidyverse, it's job is to return all rows from x where there are not matching values in y. In this case x is *the_hobbit*, y is stop_words, a tibble from the tidytext library. Also notice that now each word has it's own row in the tibble, which is now length over 30,000 elements long.

<br  />

```{r}
head(the_hobbit)
```

```{r}
head(tidy_hobbit)
```

<br  />

### Word Frequency Analysis 

Let's see some basic summaries on the hobbit text first.

```{r}
# How many words are in The Hobbit?
str_count(the_hobbit, " ") + 1

# How many sentences?
str_count(the_hobbit, "[.!?]")

# How long should the book take to read?
((str_count(the_hobbit, " ") + 1) / 200) / 60 # number of words / 200 words-per-minute / 60 minutes
```

<br  />

I chose to use the original variable the_hobbit, as tidy text doesn't retain the punctuation, and stop words still count as words. *str_count( )* counts how many times you're given variables appear within the string it's analyzing. 200 words-per-minute is a below average adult reading speed, but The Hobbit is a story best relished, so that'll do.
 
<br  />

#### Frequency Plot
```{r}
tidy_hobbit %>%
  count(word, sort = TRUE) %>%          # count words
  filter(n > 80) %>%                    # only show words that have been used 80+ times
  mutate(word = reorder(word, n)) %>%   # sorts words by decending occurance
  ggplot(aes(word, n)) +                # plot the words vs number of occurances
    geom_col(fill = "darkslateblue") +  # blue bars
  xlab("Word") +                        # labels
  ylab("Number of Occurances") + 
  labs(title = "Most Common Meaningful Words in 'The Hobbit'") +
  coord_flip() +                        # sideways box-plot for aestetic purposes
  theme_minimal()                       # I don't like the gray default theme
```

The top five most common words in *The Hobbit* are Bilbo, dwarves, Thorin, time, and Gandalf. The only non-character word is time, but I bet the word time is pretty common in most novels. I'm suprised to see the word don't on the list, I would have classified it as a stop word. There isn't much to be suprised about on this list if you've read the book. If you haven't read the book, the list gives you a pretty good idea of what to expect!

<br  />

#### Word Cloud
```{r}
hobbit_cloud <- tidy_hobbit %>%
  count(word) %>%                                      # provide a count of each word
  with(wordcloud(word, n, max.words = 100,             # using the wordcloud library, plot word, according to n (number of count), with 100 maximum words)
                 color = brewer.pal(n = 8,"Set1")))    # use RColorBrwer pallete 'Set1'
```


The top 100 words used in The Hobbit are all jumbled together in this word cloud. It's basically the same idea as the frequency plot, except it gives less information in a more artistic way. I'm not sure what's with the color coding, I would have thought all 8 pallete colors would be used, not just five. The size of the words reflect their frequency though. 

<br  />

## Character Analysis

```{r}
tidy_hobbit$linenumber <- 1:nrow(tidy_hobbit)
```


Work in Progress
```{r}
main_chr <- as.tibble(c("bilbo", "thorin", "gandalf", "smaug", "gollum", "bard")) %>%
  unnest_tokens(word, value)

colnames(main_chr) <- "word"


main_chr_num <- tidy_hobbit %>% 
  semi_join(main_chr, "word") %>%
  count(word, sort = TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) + 
    geom_col(fill = "darkslateblue")
main_chr_num
```


```{r}
main_chr_pres <- tidy_hobbit %>% 
  semi_join(main_chr, "word") %>%
  ggplot() + 
  geom_point(
    aes(word, linenumber)
  ) + 
  labs(title = "Main Character Apperances") +
  xlab("Character") + 
  ylab("")+
  coord_flip() + 
  theme_minimal() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
main_chr_pres
  
```




tidy_hobbit %>%
  count(word, sort = TRUE) %>%          # count words
  filter(n > 80) %>%                    # only show words that have been used 80+ times
  mutate(word = reorder(word, n)) %>%   # sorts words by decending occurance
  ggplot(aes(word, n)) +                # plot the words vs number of occurances
    geom_col(fill = "darkslateblue") +  # blue bars
  xlab("Word") +                        # labels
  ylab("Number of Occurances") + 
  labs(title = "Most Common Meaningful Words in 'The Hobbit'") +
  coord_flip() +                        # sideways box-plot for aestetic purposes
  theme_minimal()                       # I don't like the gray default theme





# Frequency of joy throught the book
joy_meter <- tidy_hobbit %>% 
  inner_join(get_sentiments("nrc")) %>% 
  filter(sentiment == "joy") %>%
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = n()) %>% 
  mutate(method = "NRC")

```


bilbo_count <- str_count(the_hobbit, "Bilbo")
bilbo_count

thorin_count <- str_count(the_hobbit, "Thorin")
thorin_count

gandalf_count <- str_count(the_hobbit, "Gandalf")
gandalf_count

smaug_count <- str_count(the_hobbit, "Smaug")
smaug_count

gollum_count <- str_count(the_hobbit, "Gollum")
gollum_count

bard_count <- str_count(the_hobbit, "Bard")
bard_count
```


Work in Progress
```{r, eval=FALSE}
# The 13 Dwarves
dwalin <- sum(str_count(tidy_hobbit$word, "dwalin"))
balin <- sum(str_count(tidy_hobbit$word, "balin"))
kili <- sum(str_count(tidy_hobbit$word, "kili"))
fili <- sum(str_count(tidy_hobbit$word, "fili"))
dori <- sum(str_count(tidy_hobbit$word, "dori"))
nori <- sum(str_count(tidy_hobbit$word, "nori"))
ori <- sum(str_count(tidy_hobbit$word, "ori"))
oin <- sum(str_count(tidy_hobbit$word, "oin"))
gloin <- sum(str_count(tidy_hobbit$word, "gloin"))
bifur <- sum(str_count(tidy_hobbit$word, "bifur"))
bofur <- sum(str_count(tidy_hobbit$word, "bofur"))
bombur <- sum(str_count(tidy_hobbit$word, "bombur"))
thorin <- sum(str_count(tidy_hobbit$word, "thorin"))

# plot dwarf mentions most popular to least popular

#dwarves <- as.tibble(c(dwalin, balin, kili, fili, dori,
#             nori, ori, oin, gloin, bifur,
#             bofur, bombur, thorin))

#dwarves$names <- c("dwalin", "balin", "kili", "fili", "dori",
#                   "nori", "ori", "oin", "gloin", "bifur",
#                   "bofur", "bombur", "thorin")
#dwarves
```


<br  />

## Sentiment Analysis

```{r, eval=FALSE}
## Adjective Analysis

# Joyful Words 
nrcjoy <- get_sentiments("nrc") %>%    # from the tidytext
  filter(sentiment == "joy")

happy_cloud <- tidy_hobbit %>%
  inner_join(nrcjoy) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100,
                 color = brewer.pal(n = 8,"Dark2")))
happy_cloud
```

```{r, eval=FALSE}
# Frequency of joy throught the book
joy_meter <- tidy_hobbit %>% 
  inner_join(get_sentiments("nrc")) %>% 
  filter(sentiment == "joy") %>%
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = n()) %>% 
  mutate(method = "NRC")

joy_meter %>%
  ggplot(aes(index, sentiment)) +
  geom_col(show.legend = FALSE,
           fill = "orange") + 
  xlab("Book Progression") + 
  ylab("Joy Meter") + 
  labs(title = "Words with Joyous Sentiment in 'The Hobbit'") +
  theme_minimal()

# Sad Words 
nrcsad <- get_sentiments("nrc") %>%    # from the tidytext
  filter(sentiment == "sadness")

sad_cloud <- tidy_hobbit %>%
  inner_join(nrcsad) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100,
                 color = brewer.pal(n = 8,"Dark2")))
sad_cloud


# Frequency of sadness throught the book
sad_meter <- tidy_hobbit %>% 
  inner_join(get_sentiments("nrc")) %>% 
  filter(sentiment == "sadness") %>%
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = n()) %>% 
  mutate(method = "NRC")

sad_meter %>%
  ggplot(aes(index, sentiment)) +
  geom_col(show.legend = FALSE, 
           fill = "lightsteelblue") + 
  xlab("Book Progression") + 
  ylab("Sad Meter") + 
  labs(title = "Words with Sad Sentiment in 'The Hobbit'") +
  theme_minimal() 

  
## Fearful Words
nrcfear <- get_sentiments("nrc") %>%   
  filter(sentiment == "fear")

fear_cloud <- tidy_hobbit %>%
  inner_join(nrcfear) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100,
                 color = brewer.pal(n = 8,"Dark2")))


# Frequency of sadness throught the book
fear_meter <- tidy_hobbit %>% 
  inner_join(get_sentiments("nrc")) %>% 
  filter(sentiment == "fear") %>%
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = n()) %>% 
  mutate(method = "NRC")

fear_meter %>%
  ggplot(aes(index, sentiment)) +
  geom_col(show.legend = FALSE, 
           fill = "Red4") + 
  xlab("Book Progression") + 
  ylab("Fear Meter") + 
  labs(title = "Words with Fear Sentiment in 'The Hobbit'") +
  theme_minimal()


### Happy, Sad, and Fear all together now
ggplot() + 
  geom_smooth(aes(fear_meter$index,
                  fear_meter$sentiment, 
                  color = "Fear"),
              se = FALSE) + 
  geom_smooth(aes(joy_meter$index, 
                  joy_meter$sentiment, 
                  color = "Joy"),
              se = FALSE) + 
  geom_smooth(aes(sad_meter$index, 
                  sad_meter$sentiment, 
                  color = "Sadness"),
              se = FALSE) + 
  xlab("Book Progression Index") + 
  ylab("Sentiment Index") +
  labs(
    title = "Specific Sentiment in 'The Hobbit'"
  ) + 
  theme_minimal() + 
  scale_color_manual(
    name = "Sentiment",
    values = c("red4", "orange", "lightsteelblue")
  )
```

```{r, eval=FALSE}
# Overall Sentiment
hobbit_sent <- tidy_hobbit %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 500) %>% 
  summarise(sentiment = sum(score)) %>% 
  mutate(method = "AFINN")

hobbit_sent[["sign"]] = ifelse(hobbit_sent[["sentiment"]] >=0, "positive", "negative")

hobbit_sent %>%
  ggplot() + 
  geom_col(aes(index, 
               sentiment,
               fill = sign)
           ) + 
  scale_fill_manual(name = "Sentiment",
                    values = c("positive" = "royalblue4", "negative" = "red4"),
                    guide = guide_legend(reverse=TRUE)) + 
  xlab("Book Progression Index") + 
  ylab("Sentiment Index") + 
  ggtitle("General AFINN Method Sentiment in 'The Hobbit'") + 
  theme_minimal()
```

















